### 2025-06-24 motivationletter/motivationletter.qmd
Initial work on the motivation letter

---
author: Jayani P. Gamage
qualifications: B.Sc. (Hons) in Statistics
position: PhD student
www: jayanilakshika.netlify.app/
email: jayani.piyadigamage@monash.edu
phone: +61 4 5229 8443
#signature: sigfile.png
address:
  - Editors
  - Journal of Computational and Graphical Statistics
opening: "Dear Professor Yuguo Chen and Professor Laura M. Sangalli,"
closing: "Sincerely"
linestretch: 1.4
format: letter-pdf
---

\bigskip
Please consider our manuscript "Looking at Non-Linear Dimension Reductions as Models in the Data Space" for publication in the Journal of Computational and Graphical Statistics.

Thank you for the consideration of this manuscript. We believe that it is a good fit for the Journal
of Computational and Graphical Statistics.