### 2025-08-19 JCGS_response_to_reviews/response_to_reviews.qmd
Start working on the reviews

---
title: "Response to Reviews: JCGS-25-420"
author: Jayani P. Gamage, Dianne Cook, Paul Harrison, Michael Lydeamore
format: pdf
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```